---
title: "edsamplr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{edsamplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edsamplr)
```

Edsamplr is an R package that makes it easier to get data (through simulation) for teaching statistics. Below is an overview of some of the things edsamplr can do. Edsamplr was created by Keah Tandon and Michael A. Seaman.

## Why simulate data?



\

## Types of data

The current version of edsamplr (add number) has been designed around the concepts covered in a first-semester statistics course. As such, the data that can be simulated includes:

- *p* variables of binary data

- *p* variables of categorical data with *k* groups

- *k* groups of continuous data (with optional specified standardized effect size)

- 2 variables of correlated continuous data with specified slope for 1 group

- Matched pairs continuous data with specified correlation (and optional specified standardized effect size) for 1 group

- Heteroscedastic continuous data with specified degree of heteroscedasticity for 1 group

- Named continuous distributions:

  - Beta
  
  - Cauchy
  
  - Chi-squared
  
  - Exponential
  
  - F
  
  - Gamma
  
  - Laplace
  
  - Logistic
  
  - Log normal
  
  - Normal
  
  - T
  
  - Triangular
  
  - Uniform
  
  - Weibull
  
\

## How to use

1. Identify a concept.

  First, identify the concept that you want to teach and determine what type of data you need for teaching that concept.

2. Decide what numbers to use as input parameters.

  If you have found a published article that demonstrates this concept but doesn't have the associated data available, you can use the sample statistics that the authors provided as parameters to simulate from. Your resultant data will not match the article's statistics exactly, but it will be pretty close.

  If you haven't found an article that demonstrates the concept (or don't have time to look), you can select any numbers to use as parameters to simulate data from. We recommend developing a short scenario to go along with the data so that your students can better connect to the data and make meaning from the lesson.

3. Simulate data.

  Use the function that will simulate the data that you need for your concept. Use the input parameters that you decided upon in the previous step in the appropriate argument.
  
  The functions contain optional summary table arguments. If you enable this, the output will contain the input parameters and summary statistics for the sample that was simulated to show how closely the sample resembles the parameters. If it is not close enough for your liking, you can re-run the code to simulate a new sample.
  
  The functions also have an optional argument for naming the variable(s). You can use this to pre-name the output columns according to your article/scenario.

4. Present to students.

  Whether your plan is to use this in class as part of your lesson, outside of class for a homework assignment, or as part of an assessment, you're ready to share with your students. If you based the data sample on a published article, make sure to clarify for your students that this is data sampled with the article's statistics as parameters and is not the original data.

\

## Additional resources

To learn more about edsamplr, please read the original edsamplr paper [add link]. Questions about edsamplr can be sent to Keah Tandon (tandonk@mailbox.sc.edu). Bugs and feature requests should be submitted to the edsamplr development page [add link] on GitHub.


For specific examples, check out the other vignettes.
