---
title: "Simulating Proportions Data"
output: rmarkdown::html_vignette
bibliography: Proportions.bib
csl: apa
vignette: >
  %\VignetteIndexEntry{Proportions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends: tidyverse
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
require(tidyverse)

devtools::load_all()

```

``` {r}
# library(edsamplr)
# library(tidyverse)

```


### Introduction

Imagine you are teaching your students about proportions. You want to give your students a real-life example from published research, so you search around and find an experimental article on scopaesthesia [@sheldrake2024].

Unfortunately, you cannot find the original data and you do not have have time to contact the authors. Fortunately, you can use the statistics from the article as parameters for simulating sample data.

### Article Abstract

Scopaesthesia, also known as the sense of being stared at, is widely reported and supported by controlled experiments showing detection rates significantly above chance when blindfolded subjects are observed directly by starers who are physically present. By contrast, indirect observation by remote starers through CCTV has demonstrated effects detectable physiologically by changes in electrodermal activity, but which do not reach conscious awareness. Further to explore indirect scopaesthesia, we tested whether people could tell when they were being watched on computer or phone screens. Participants worked in pairs. In each randomized trial, the starer either saw a live video of the staree, or a 'distraction' photograph. Starees then reported if they felt stared at or not. We analysed our results statistically for the overall hit rate, and for variables such as feedback, observer-subject relationship, and distance. Across 6,050 trials, the overall hit rate was 49.9%, almost exactly at the 50% chance level. The chance-level outcome was not affected by feedback, gender of starers and starees, distance between starers and starees, or multiple starers compared with single starers. Thus, scopaesthesia was undetectable consciously when people were stared at indirectly via screens. [@sheldrake2024]

### Selected Findings to Sample From

Sheldrake and Penraeth [-@sheldrake2024] had a much more complex design than you need for your class, so you decide to focus on the hit proportion for the 'looking' condition.

\

**Explanatory variable**: 'Looking' condition

**Response variable**: Successful identification of staring

```{r}
staring_outcomes <- data.frame(Trials = c(2914, 2926),
                               Hits = c(1654, 1255),
                               Hit_Rate = c(56.8, 42.9),
                               row.names = c("Seen", "Unseen"))

staring_outcomes

```

You can use the `generate_categorical()` function to simulate data that will resemble this.

We have two variables with two conditions each, so this is what our frequency table looks like:

```{r}
staring_counts <- data.frame(Hits = staring_outcomes$Hits,
                             Misses = staring_outcomes$Trials - staring_outcomes$Hits,
                             row.names = c("Seen", "Unseen"))

staring_counts


```

We can turn that into a proportion table.

``` {r}
staring_props <- round(staring_counts/sum(staring_counts), 3)

staring_props

```

In this situation, we want to get a sample of data that matches this proportion table. It may seem counter intuitive to run generate_categorical() with one variable and 4 groups, but we are not trying to simulate two separate categorical variables, rather 4 cross-conditions.

I like to define my arguments prior to calling the function so that I can make sure everything lines up. When using the separate argument, make sure that you put the explanatory variable first, then the result variable.

```{r}
p <- 1
n <- 5840
k <- 4
k_prop <- c(0.283, 0.216, 
            0.215, 0.286)
k_names <- c("Seen_Hit", "Seen_Miss", 
           "Unseen_Hit", "Unseen_Miss")

output <- generate_categorical(p = p, 
                               n = n, 
                               k = k, 
                               k_prop = k_prop,
                               k_names = k_names,
                               separate = TRUE)

```

Our output provides both the simulated data and a descriptive summary comparing the inputs to the simulated data.

``` {r}
data <- output$sample

output$proportion_summary

```

If you're happy with the final product, you're ready to export the data for your students. If you'd like to try to get it closer to the original data, you can rerun the function.

### References
