---
title: "Teaching Independent Samples t Test"
output: 
  rmarkdown::html_vignette:
    csl: apa.csl
bibliography: t_test.bib
resource_files:
  - apa.csl
vignette: >
  %\VignetteIndexEntry{Teaching Independent Samples t Test}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends: tidyverse
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(edsamplr)
require(tidyverse)

```

```{r setup}
# library(edsamplr)
# library(tidyverse)
```

### Introduction

Imagine you want to teach your students about independent samples $t$ tests. You want to give your students a real-life example, and you recently came across Deng and McShane's [-@deng2025] article on emojis in marketing. Unfortunately, you cannot find the original data and you do not have have time to contact the authors. Fortunately, you can use the statistics from the article as parameters for simulating sample data.

### Article Abstract

Contributing to a burgeoning area of research on the nuanced effects of emojis in brand communications, the current research builds understanding of two dominant forms of emoji role—emojis as text reinforcement and emojis as text substitution—and their downstream effects. Across three studies, we examine how emoji roles differentially interact with message features to influence brand-level outcomes (brand attitudes, product quality expectations and consumers' willingness to try a brand's product) through their effects on processing fluency. We find robust evidence that substitution emojis elicit more negative brand-level outcomes than reinforcement emojis both when the emoji has low congruence with the text and when the complexity of the text in the message is high, and that these effects are mediated by processing fluency. These findings deepen our understanding of emojis' effects in brand communications and provide practical guidance for digital marketers regarding how to effectively leverage emojis. [@deng2025]

### Selected Findings to Sample From

Deng and McShane [-@deng2025] had multiple studies and a more complex design, but their first analysis in their first study was an independent samples $t$ test. The authors compared processing fluency in tweets that contained emojis and text and those that just contained text and found that there was a statistically significant difference between the two groups [@deng2025]. $$M_{emoji} = 5.86 \text{ vs. } M_{text} = 6.27, t(247) = 2.73, p = 0.003$$

\

**Explanatory variable**: Tweet type

**Response variable**: Processing fluency

\

Since we don't have the variance or standard deviation of the two groups, we will have to calculate that from what we do have. We know from the $F$ tests later in the article that the sample sizes are $n_{emoji}=171$ and $n_{text}=78$, and we can use the results from those $F$ tests to calculate the variances.

``` {r}

# what we know

n_emoji <- 171
n_text <- 78
m_emoji <- 5.86
m_text <- 6.27
t <- 2.73

# calculating variance from F stats

m_emoji_r <- 6
m_emoji_s <- 5.73
n_emoji_r <- 87
n_emoji_s <- 84
F_rs <- 2.15

diff_rs <- m_emoji_r - m_emoji_s
var_emoji <- diff_rs^2/(F_rs * (1/n_emoji_r + 1/n_emoji_s))

F_st <- 9.65

se2_total <- (m_text - m_emoji_s)^2/F_st
var_text <- (se2_total - (var_emoji/n_emoji_s)) * n_text

```

Now that we have the variances, we can simulate data. The `generate_quantitative()` function provides us the sample data and summary statistics comparing our input to our output. We can't get an exact match to the original data, but we can get close.

``` {r}

sample <- generate_quantitative(k = 2,
                              mean = c(m_emoji, m_text),
                              var = c(var_emoji, var_text),
                              n = c(n_emoji, n_text),
                              k_names = c("emoji", "text"))

data <- sample$sample

sample$summary

```

Just to check to see, here is a $t$ with the resulting data. We were not able to recreate their findings, but we had similar data.

```{r}

t.test(Value ~ Group, data = data, var.equal = TRUE)

```

If you're happy with the final product, you're ready to export the data for your students. If you'd like to try to get it closer to the original findings, you can update the arguments in the function and re-run. Either way, make sure to tell your students that this is not the original data but a sample simulated from moments that are similar to the statistics from the original.

### References
